### 2016-07-06 FS4/GSS/gss_w6.tex
6.5.2

% import math packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
% contradiction lightning
\usepackage{stmaryrd}
% algorithms and pseudo code
\usepackage{algorithmic}
\usepackage{algorithm}
%\usepackage{clrscode3e}
% formatting and layout
\usepackage{color}
% settings
\usepackage{footnote}
\usepackage{perpage}
\MakePerPage{footnote}
% custom text colors
\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}
% header and footer
\usepackage{fancyhdr}
% filler text
\usepackage{lipsum}
% macro commands
%% requires color package and the custom colors defined here 
\newcommand{\todo}[1]{
	\addcontentsline{toc}{subsubsection}{TODO: #1}
	\textcolor{pgreen}{\texttt{\-\\ \-\\//TODO: #1\-\\ \-\\}}
}
\newcommand{\sect}[3]{	%	custom section (level 1)
	\newpage
	\addcontentsline{toc}{section}{Zettel #1 (#2)}
	\section*{Zettel Nr. #1 (Ausgabe: #2, Abgabe: #3)}
	\label{sec:#1}
}
\newcommand{\ttt}[1]{\texttt{#1}}

@@

										Ebenso kann er nicht selber Tickets editieren oder bearbeiten.	\\
			\textbf{Verhalten}		&	aktiv: liest veränderte Tickets am Automaten ein	\\
			\textbf{Ressourcen}		&	beschränkt: nicht genügend Rechenkapazität, um den Algorithmus zu knacken;
										Er hat außer über die Barcodescanner keinen Zugriff auf das System, insbesondere stehen ihm keine weiteren Schnittstellen zur Verfügung
		\end{tabular}
	\end{savenotes}
\end{center}
\subsubsection{Umsetzung mit kryptographischen Techniken (Pflicht; 4 Punkte)}

@@

\stepcounter{subsection}
\subsection{RSA-Verfahren}
\stepcounter{subsubsection}
\subsubsection{Anwendung (Pflicht; 6 Punkte)}

>>>>

% import math packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
% contradiction lightning
\usepackage{stmaryrd}
% algorithms and pseudo code
\usepackage{algorithmic}
\usepackage{algorithm}
%\usepackage{clrscode3e}
% formatting and layout
\usepackage{color}
% settings
\usepackage{footnote}
\usepackage{perpage}
\MakePerPage{footnote}
% custom text colors
\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}
\definecolor{light-light-gray}{gray}{0.95}
% header and footer
\usepackage{fancyhdr}
% filler text
\usepackage{lipsum}
% source code inclusion
\usepackage{listings}
\lstset{ %
language=Java,   							% choose the language of the code
basicstyle=\small\ttfamily,  				% the size of the fonts that are used for the code
numbers=left,                   			% where to put the line-numbers
numbersep=5pt,                  			% how far the line-numbers are from the code
backgroundcolor=\color{light-light-gray},   % choose the background color. You must add
frame=lrtb,           						% adds a frame around the code
tabsize=4,          						% sets default tabsize to 2 spaces
captionpos=t,           					% sets the caption-position to top
breaklines=true,        					% sets automatic line breaking
breakatwhitespace=true,
xleftmargin=1.5cm,							% space from the left paper edge
showspaces=false,
showstringspaces=false,
showtabs=false,
commentstyle=\color{pgreen},
keywordstyle=\color{pblue},
literate=%
    {Ö}{{\"O}}1
    {Ä}{{\"A}}1
    {Ü}{{\"U}}1
    {ß}{{\ss}}1
    {ü}{{\"u}}1
    {ä}{{\"a}}1
    {ö}{{\"o}}1
    {~}{{\textasciitilde}}1
}
\renewcommand{\lstlistingname}{Code}
% macro commands
%% requires color package and the custom colors defined here 
\newcommand{\todo}[1]{
	\addcontentsline{toc}{subsubsection}{TODO: #1}
	\textcolor{pgreen}{\texttt{\-\\ \-\\//TODO: #1\-\\ \-\\}}
}
\newcommand{\sect}[3]{	%	custom section (level 1)
	\newpage
	\addcontentsline{toc}{section}{Zettel #1 (#2)}
	\section*{Zettel Nr. #1 (Ausgabe: #2, Abgabe: #3)}
	\label{sec:#1}
}
\newcommand{\ttt}[1]{\texttt{#1}}

@@

										Ebenso kann er nicht selber Tickets editieren oder bearbeiten.	\\
			\textbf{Verhalten}		&	aktiv: liest veränderte Tickets am Automaten ein	\\
			\textbf{Ressourcen}		&	beschränkt: nicht genügend Rechenkapazität, um den Algorithmus zu knacken;
										Er hat außer über die Barcodescanner keinen Zugriff auf das System, insbesondere stehen ihm keine weiteren Schnittstellen zur Verfügung

										Wir gehen davon aus, dass die Ressourcen derart beschränkt sind, dass der Angreifer nicht in der Lage ist, durch Heuristiken, Kombinatorik, Inferenz oder Brute Force das Verfahren der Code-Generierung herauszufinden.
		\end{tabular}
	\end{savenotes}
\end{center}
\subsubsection{Umsetzung mit kryptographischen Techniken (Pflicht; 4 Punkte)}

@@

\stepcounter{subsection}
\subsection{RSA-Verfahren}
\stepcounter{subsubsection}
\subsubsection{Anwendung (Pflicht; 6 Punkte)}
Folgender Python-Code berechnet die für RSA notwendigen Werte, liest die verschlüsselte Nachricht ein und speichert die entschlüsselte Nachricht in einem Textfile:
\begin{lstlisting}[caption=\string$ python rsa.py msg.txt]
from sys import argv  # to have access to console parameters

def extended_gcd(aa, bb):
    """
    SOURCE: http://rosettacode.org/wiki/Modular_inverse#Python
    returns the modular multiplicative inverse of a number aa given to a base bb using the extended euclidean algorithm
    :param aa: the number to invert
    :param bb: the base
    :return: the modular inverse to [aa]bb
    """
    lastremainder, remainder = abs(aa), abs(bb)
    x, lastx, y, lasty = 0, 1, 1, 0
    while remainder:
        lastremainder, (quotient, remainder) = remainder, divmod(lastremainder, remainder)
        x, lastx = lastx - quotient * x, x
        y, lasty = lasty - quotient * y, y
    return lastremainder, lastx * (-1 if aa < 0 else 1), lasty * (-1 if bb < 0 else 1)

def modinv(a, m):
    """
    SOURCE: http://rosettacode.org/wiki/Modular_inverse#Python
    returns the modular inverse of a given number to a given base, if there is one
    :param a: the number to invert
    :param m: the base
    :return: the modular inverse to [a]m, if there is one, ValueError else
    """
    g, x, y = extended_gcd(a, m)
    if g != 1:
        raise ValueError
    return x % m

def getwords(file):
    """
    SOURCE: http://stackoverflow.com/questions/10264460/read-the-next-word-in-a- file-in-python (accepted answer)
    This function was slightly modified compared to the source
    """
    with open(file) as wordfile:
        wordgen = words(wordfile)
        single_words = []
        for word in wordgen:
            if word.startswith(" "):
                word = word[1:]
            if len(word) > 0 and word != '\n':
                single_words.append(word)
        return single_words

def words(fileobj):
    """
    SOURCE: http://stackoverflow.com/questions/10264460/read-the-next-word-in-a- file-in-python (accepted answer)
    This function was slightly modified compared to the source
    """
    for line in fileobj:
        for word in line.split(","):
            yield word

# read encrypted message into RAM (assuming the file contains nothing but the encrypted message)
words = getwords(argv[1])
# get encryption details
p = 271
q = 379
e = 47
phi = (p - 1) * (q - 1)

# calculate decryption exponent
d = modinv(e, phi)
# casts the results, which are 'long', as 'int'
res = map(lambda x: int(x),
          # equivalent to: 'for all numbers in $words, raise them to $d and modulate them with $phi
          map(lambda x: (int(x) ** d) % phi,
              words))

# for debugging purposes, show the decrypted numbers in the console
print res

# store the result in a text file
out = open('(%s)-dec.txt' % argv[1], 'w')
for e in res:
    out.write(str(e))
    out.write(", ") # separate each number by a comma and a whitespace
out.close()

\end{lstlisting}
Die entschlüsselten Zahlen sehen wie folgt\footnote{\textit{Anmerkung:} Es wurden manuell Zeilenumbrüche eingefügt, die in der eigentlichen Ausgabedatei nicht enthalten waren} aus (nach Code):
\begin{lstlisting}[caption=(msg.txt)-dec.txt]
80919, 49559, 79877, 2603, 6073, 40095, 43740, 79877, 6073, 7469, 46204,
46204, 88109, 67068, 49708, 85888, 49559, 71999, 49559, 2603, 6073, 71999,
43740, 16540, 40095, 6073, 83029, 28431, 49708, 39320, 79877, 16540, 40095,
79877, 6073, 81236, 98756, 79877, 37927, 79877, 16540, 6073, 94151, 43740,
58492, 98756, 98797, 43740, 39320, 43740, 6073, 72071, 16540, 39320, 2603,
79877, 43740, 83029, 79877, 2603, 37927, 28431, 40095, 79877, 49708, 49708,
79877, 61965, 6073, 46204, 58492, 98756, 49559, 98797, 20896, 20896, 43740,
79877, 49708, 79877, 61965, 6073, 1820, 85888, 43740, 16540, 29889, 28431,
94151, 6073, 81236, 85888, 29889, 49708, 79877, 71999, 61965, 6073, 40095,
43740, 79877, 6073, 43447, 96957, 16540, 88109, 22259, 46204, 43740, 58492,
98756, 79877, 2603, 98756, 79877, 43740, 98797, 6073, 163, 28431, 16540,
6073, 52488, 85888, 71999, 71999, 94151, 28431, 79877, 2603, 98797, 79877,
2603, 16540, 6073, 49559, 16540, 40095, 6073, 40095, 85888, 20896, 49559,
39320, 79877, 98756, 28431, 79877, 2603, 43740, 39320, 79877, 6073, 72071,
16540, 39320, 2603, 43740, 83029, 83029, 79877, 61965, 6073, 55061, 49559,
39320, 85888, 16540, 39320, 71999, 88109, 6073, 49559, 16540, 40095, 6073,
55061, 49559, 39320, 2603, 43740, 83029, 83029, 71999, 45927, 28431, 16540,
98797, 2603, 28431, 49708, 49708, 79877, 61965, 6073, 81236, 43740, 37927,
43740, 16540, 39320, 88109, 72071, 98797, 98797, 85888, 58492, 45927, 6073,
49559, 16540, 40095, 6073, 52488, 28431, 94151, 79877, 2603, 88109, 72071,
16540, 85888, 49708, 82007, 71999, 43740, 71999, 61965, 6073, 78732, 43740,
28431, 37927, 79877, 98797, 2603, 43740, 71999, 58492, 98756, 79877, 6073,
26188, 79877, 2603, 83029, 85888, 98756, 2603, 79877, 16540, 61965, 6073,
7469, 2603, 49559, 16540, 40095, 49708, 85888, 39320, 79877, 16540, 6073,
40095, 79877, 2603, 6073, 67068, 2603, 82007, 92583, 98797, 28431, 39320,
2603, 85888, 92583, 98756, 43740, 79877, 61965, 6073, 40095, 85888, 71999,
6073, 1820, 46204, 72071, 88109, 26188, 79877, 2603, 83029, 85888, 98756,
2603, 79877, 16540, 61965, 6073, 72071, 49559, 98797, 98756, 79877, 16540,
98797, 43740, 83029, 43740, 45927, 85888, 98797, 43740, 28431, 16540, 71999,
92583, 2603, 28431, 98797, 28431, 45927, 28431, 49708, 49708, 79877, 6073,
49559, 16540, 40095, 6073, 16540, 85888, 98797, 49559, 79877, 2603, 49708,
43740, 58492, 98756, 6073, 85888, 49708, 49708, 79877, 6073, 85888, 16540,
40095, 79877, 2603, 79877, 16540, 6073, 24057, 16540, 98756, 85888, 49708,
98797, 79877, 61965, 6073, 40095, 43740, 79877, 6073, 94151, 43740, 2603,
6073, 43740, 16540, 6073, 40095, 79877, 2603, 6073, 96957, 79877, 29889,
49559, 16540, 39320, 6073, 49559, 16540, 40095, 6073, 40095, 79877, 2603,
6073, 26188, 28431, 2603, 49708, 79877, 71999, 49559, 16540, 39320, 6073,
29889, 79877, 98756, 85888, 16540, 40095, 79877, 49708, 98797, 6073, 98756,
85888, 29889, 79877, 16540, 6073, 43740, 88109, 22259, 
\end{lstlisting}
